---
title: "Darwin Core Mapping UTM SQLite connection"
subtitle: "VBP-INBO-waarnemingen-occurrences_2026_gewervelden_hoge resolutie
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
library(ggplot2)        # for map
library(mapview)        # to view map
library(DBI)            # database stuff
library(RSQLite)        # sqlite connection
library(reader)
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

#load and list files


# Verbind met SQLite

```{r eval=FALSE, include=FALSE}
con <- dbConnect(SQLite(), (here::here("observation-2026-01-29-15-46-observations-instituut-voor-natuur-en-bosonderzoek.sqlite" )))
```


# Bekijk alle tabellen én views

```{r eval=FALSE, include=FALSE}
dbListTables(con)  # views staan hier ook tussen
```

```{r eval=FALSE, include=FALSE}
input_interim <- dbReadTable(con, "INBO_WNMN_CHORDATA_from2019")  # werkt ook voor views
```


# Resultaat bekijken

```{r}
head(input_interim)
```





# remove spec.

```{r}
input_interim <- input_interim %>% 
  mutate(
    scientificName = scientificName %>% 
      str_remove("\\s+spec\\.?$") %>%  # verwijdert ' spec.' of ' spec'
      str_trim()
  )
```



Preview data:


# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```

# embargo date

```{r}
input_interim <- input_interim %>%
  filter(is.na(embargo_date) | embargo_date < today())
```





# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## remove obsolete columns

```{r}
occurrence %<>% select(-c(is_escape, embargo_date))
```


# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/10.15468/xa5sqf")
```

### add row numbers to genrate occurrenceID unique


```{r}
occurrence %<>% mutate(ID_row = row_number())
```


### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "By accessing this dataset, you agree to use it solely for internal analysis. You may not reproduce, distribute, or share the data.") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "INBO") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "only for internal use") 
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_references = references) 
```



### datasetID

```{r}
#occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "INBO") # e.g. "INBO"
```
### institutionID

```{r}
occurrence %<>% mutate(dwc_institutionID = "https://ror.org/00j54wy13") # e.g. "INBO"
```
### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "INBO collaborator Occurrences -High resolution - of vertebrates recorded in waarnemingen.be in Flanders, Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

```{r}
occurrence %<>% mutate(dwc_dataGeneralizations = "coordinates are not generalized")
```



### occurrenceID

```{r}
occurrence %<>% mutate(dwc_occurrenceID = str_c("INBO:OCC:WNMN:SEN:", occurrenceID))
  
```

```{r}
occurrence %<>% mutate(dwc_catalogNumber = references)
```



### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = recordedBy)
```

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = indentifiedBy)
```



### individualCount

```{r}
occurrence %<>%  mutate(dwc_individualCount = individualCount) 
                
```

```{r}
occurrence$individualCount %<>% replace_na(0)
```

### organismQuantity


```{r}
occurrence %<>%  mutate(dwc_organismQuantity = individualCount) 
                
```



### organismQuentityType

```{r}
occurrence %<>%  mutate(dwc_organismQuantityType = 'count estimate') 
                
```


### sex

```{r}
occurrence %<>% mutate(dwc_sex = case_when(occurrence$sex == 'U' ~'unknown'
                                       ,occurrence$sex == 'F' ~'female'
                                       ,occurrence$sex == 'M' ~'male',
                                       TRUE ~ sex))
```



### lifeStage

```{r}
occurrence %<>% mutate(dwc_lifestage = case_when(occurrence$`lifeStage` == 'onbekend' ~'unknown'
                                                 , TRUE ~ `lifeStage`))
```


### behavior

```{r}
occurrence %<>% mutate(dwc_behavior = behaviour)
```

### occurrenceRemarks

```{r}
occurrence %<>% mutate(dwc_collectionCode = 'INBO-WNMN:SEN')
```

## samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingEffort = 'casual observation')
```

```{r}
occurrence %<>% rename(dwc_samplingProtocol = samplingProtocol)
 


```

### samplingEffort


### eventDate

eventDate already ok in source file (should be)

```{r}


#occurrence %<>% ymd(date)

occurrence %<>% mutate(dwc_eventDate = eventDate)
  


```

# Location

```{r}
occurrence %<>%
  mutate(dwc_decimalLongitude = decimalLongitude) %>%
  mutate(dwc_decimalLatitude = decimalLatitude) %>%
  mutate(dwc_geodeticDatum = "WGS84") %>%
  mutate(dwc_coordinateUncertaintyInMeters = "30") %>%
  mutate(dwc_stateProvince = stateProvince) %>%
  mutate(dwc_countryCode = "BE")  %>%            
  mutate(dwc_continent = "Europe")
```

```{r}

occurrence %<>%
  
  mutate(dwc_decimalLongitude = round(dwc_decimalLongitude, digits = 5)) %>%
  mutate(dwc_decimalLatitude = round(dwc_decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(dwc_decimalLatitude = as.character(format(dwc_decimalLatitude, nsmall = 5))) %>%
   mutate(dwc_decimalLongitude = as.character(format(dwc_decimalLongitude, nsmall = 5)))
```
### coordinateUncertaintyInMeters

```{r}
occurrence %<>%
  mutate(dwc_coordinateUncertaintyInMeters = '30')
```


### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
#occurrence %<>% mutate(countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
occurrence %<>% mutate(dwc_municipality = sub(" -.*","",locality))
```


```{r}
occurrence %<>%
  mutate(dwc_locality = locality)
```

### verbatimcoordinates


### verbatimLatitude

### verbatimLongitude

### verbatimcoordinatesystem

```{r}
occurrence %<>% mutate(verbatimcoordinatesystem = "WGS84") # 
```

### verbatimSRS

```{r}
#occurrence %<>% mutate(verbatimSRS = "Belgian Datum 1972")
```

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

### identifiedBy



### identificationVerificationStatus

```{r}
occurrence %<>% mutate(dwc_identificationVerificationStatus = identificationVerificationStatus)
```


## dynamicProperties

```{r}
occurrence %<>% mutate(dwc_dynamicProperties = '{"rbac":true,"rbac_allowed":"HIGHRES"}')
```


### identificationRemarks




### scientificName 

```{r}
occurrence %<>% mutate(dwc_scientificName = scientificName)
```


### vernacularName 

```{r}
occurrence %<>% mutate(dwc_vernacularName = vernacularName)
```


### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```


### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = taxonRank )
```



### nomenclaturalCode

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(dwc_occurrenceStatus = case_when(dwc_individualCount > 0 ~ "Present",
                                        dwc_individualCount == 0 ~ "Absent",
                                        is.na(dwc_individualCount) ~ "Absent"
                                                  
                                       ))
head(occurrence, n = 5)
```



## Post-processing

```{r}

occurrence %<>% select(starts_with("dwc_"))
```



```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")

```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

Preview data:



```{r}
occurrence %>% head()
```

### add function for scientificNameAuthorship (Bijgevoegd stuk!!!!!)

## unique names

```{r}
unique_names <- occurrence %>%
  distinct(scientificName) %>%
  pull(scientificName)

```
```{r}
get_backbone_info <- function(name) {
  res <- tryCatch({
    gbif_res <- name_backbone(name = name)
    tibble(
      scientificName = name,
      scientificNameAuthorship = gbif_res$authorship %||% NA,
      canonicalName = gbif_res$canonicalName %||% NA,
      taxonKey = gbif_res$usageKey %||% NA,
      taxonomicStatus = gbif_res$status %||% NA,
      matchType = gbif_res$matchType %||% NA
    )
  }, error = function(e) {
    tibble(
      scientificName = name,
      scientificNameAuthorship = NA,
      canonicalName = NA,
      taxonKey = NA,
      taxonomicStatus = NA,
      matchType = NA
    )
  })
  return(res)
}
```

```{r}
gbif_lookup <- map_dfr(unique_names, get_backbone_info)
```

```{r}
get_taxon_info <- function(taxonKey) {
  tryCatch({
    res <- name_usage(taxonKey)
    tibble(
      taxonKey = taxonKey,
      scientificName = res$data$scientificName,
      canonicalName = res$data$canonicalName,
      scientificNameAuthorship = res$data$authorship,
      rank = res$data$rank,
      status = res$data$status,
      kingdom = res$data$kingdom,
      phylum = res$data$phylum,
      class = res$data$class,
      order = res$data$order,
      family = res$data$family,
      genus = res$data$genus
    )
  }, error = function(e) {
    tibble(
      taxonKey = taxonKey,
      scientificName = NA,
      canonicalName = NA,
      scientificNameAuthorship = NA,
      rank = NA,
      status = NA,
      kingdom = NA,
      phylum = NA,
      class = NA,
      order = NA,
      family = NA,
      genus = NA
    )
  })
}
```

```{r}
gbif_info <- map_dfr(gbif_lookup$taxonKey, get_taxon_info)
```

maak GBIF_info uniek

```{r}
gbif_info_unique <- gbif_info %>%
  distinct(canonicalName, .keep_all = TRUE)
```

```{r}
duplicated_rows <- gbif_info %>%
  group_by(canonicalName) %>%
  filter(n() > 1) %>%
  ungroup()
```



# Left join: behoud alle rijen van occurrence
```{r}
occurrence_joined <- occurrence %>%
  left_join(gbif_info_unique, by = c("scientificName" = "canonicalName"))
```
```{r}
occurrence_joined <- occurrence_joined %>%
  distinct(occurrenceID, .keep_all = TRUE)
```


Save to CSV:

```{r}
write_csv(occurrence_joined, here::here("datasets", "np-datavragen-data-gewervelden", "data", "processed", "occurrence_chordata_high_res.csv"), na = "")
```


# Visualize data on a map OriginalCoördinates

```{r}
occurrenceMap <- occurrence  %>%
 filter(identifiedBy == "Yolan Bosteels")
```

```{r}
occurrenceMap$decimalLatitude <- as.numeric(occurrenceMap$decimalLatitude)
```

```{r}
occurrenceMap$decimalLongitude <- as.numeric(occurrenceMap$decimalLongitude)
```

```{r}
head(occurrenceMap)
```



```{r}


head(occurrence)

mapview(occurrenceMap, xcol = "decimalLongitude", ycol = "decimalLatitude", crs = 4269, grid = FALSE)

```

# Visualize data on a map UTMCoördinates

```{r}
occurrenceMapUTM <- occurrence %>% 
 filter(identifiedBy == "Jeroen Vanden Borre")
```

```{r}
occurrenceMapUTM$centroid_latitude <- as.numeric(occurrenceMapUTM$centroid_latitude)
```

```{r}
occurrenceMapUTM$centroid_longitude <- as.numeric(occurrenceMapUTM$centroid_longitude)
```

```{r}
head(occurrenceMapUTM)
```
```{r}
problems(input_interim)
```



```{r}


head(occurrence)

mapview(occurrenceMapUTM, xcol = "centroid_longitude", ycol = "centroid_latitude", crs = 4269, grid = FALSE)

```

