---
title: "Darwin Core Mapping"
subtitle: "VBP-GBIF data downloaded from various wnmn profiel -fork mica W-Vl ~species group bijen wespen en mieren"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
library(ggplot2)        # for map
library(mapview)
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

#load and list files


```{r}
csv_files <- list.files(here::here("datasets", "np-profile-download", "data", "raw"), pattern = "*.csv$", full.names = TRUE)

```


# read all csv files

```{r eval=FALSE, include=FALSE}
csv_files <- list.files(here("datasets", "np-profile-download", "data", "raw"), 
                        pattern = "\\.csv$", full.names = TRUE)
```


#read and add filename (old-no run)

```{r eval=FALSE, include=FALSE}
read_and_add_filename <- function(file) {
  observerN <- read_csv(file)
  observerN$filename <- basename(file)  # Add filename as column in dataframe
  observerN
}
```



# Function to read the file and add the relevant part of the filename (using ReGex)
```{r}
read_and_add_filename <- function(file) {
  df <- read_csv(file)
  # Bestandsnaam zonder pad
  filename <- basename(file)
  
  # Deel na 'observations-' tot aan het volgende '-'
  extracted_part <- str_extract(filename, "(?<=observations-)([^\\-]+)-[^\\-]+") %>% 
  gsub("-", " ", .) %>% gsub(".csv", " ",.)
  
  # Toevoegen aan dataframe
  df$Extracted <- extracted_part
  df
}
  
```

  
 

# Data inlezen en samenvoegen
```{r}
data <- map_dfr(csv_files, read_and_add_filename)

```

```{r eval=FALSE, include=FALSE}
data <- lapply(csv_files, read_csv) %>% bind_rows()
```

# Resultaat bekijken

```{r}
head(data)
```





```{r}
#data <- bind_rows(csv_files)
```

```{r}
input_interim <- data
```



Preview data:

```{r}
input_interim %>% head(n = 5)
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```

# Georeference source data

## Lambert to Decimals

### Keep original x,Y as verbatimcoordinates



# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## remove obsolete columns

```{r}
occurrence %<>% select(-c(modified))
```
## filter for species group

Define which group of species to map to DwC

```{r}
occurrence %<>%
        filter(`species group` == 'Bijen, wespen en mieren') 
        
head(occurrence, n = 5 )

```



# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/10.15468/XXXXX")
```

### add row numbers to genrate occurrenceID unique


```{r}
occurrence %<>% mutate(ID_row = row_number())
```


### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = str_c(str_to_title(Extracted) )) # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_references = link) 
```


### datasetID

```{r}
#occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "INBO") # e.g. "INBO"
```

### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "VBP - Citizen science Occurrences of Bees, wasps, and ants (Apidae, Vespidae, Sphecidae and Formicidae in  Flanders, Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

### occurrenceID

```{r}
occurrence %<>% mutate(dwc_occurrenceID = str_c("VBP:OCC:", source, id , ":", link))
  
```

### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = Extracted)
```

### individualCount

```{r}
occurrence %<>%  mutate(dwc_individualCount = number) 
                
```

```{r}
occurrence$individualCount %<>% replace_na(0)
```

### organismQuantity


```{r}
occurrence %<>%  mutate(dwc_organismQuantity = number) 
                
```



### organismQuentityType

```{r}
occurrence %<>%  mutate(dwc_organismQuantityType = 'individuals') 
                
```


### sex

```{r}
occurrence %<>% mutate(dwc_sex = case_when(occurrence$sex == 'u' ~'unknown'
                                       ,occurrence$sex == 'F' ~'female'
                                       ,occurrence$sex == 'M' ~'male',
                                       TRUE ~ sex))
```



### lifeStage

```{r}
occurrence %<>% mutate(dwc_lifestage = case_when(occurrence$`life stage` == 'onbekend' ~'unknown'
                                                 , TRUE ~ `life stage`))
```


### behavior

### occurrenceRemarks

```{r}
occurrence %<>% rename(dwc_occurrenceRemarks = notes)
```

## samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingEffort = 'casual observation')
```

```{r}
occurrence %<>% rename(dwc_samplingProtocol = `counting method`)
 


```

### samplingEffort


### eventDate

eventDate already ok in source file (should be)

```{r}


#occurrence %<>% ymd(date)

occurrence %<>% mutate(dwc_eventDate = date)
  


```

# Location

```{r}
occurrence %<>%
  mutate(dwc_decimalLongitude = lng) %>%
  mutate(dwc_decimalLatitude = lat) %>%
  mutate(dwc_geodeticDatum = "WGS84") %>%
  mutate(dwc_coordinateUncertaintyInMeters = "30") %>%
  mutate(dwc_stateProvince = `country division`) %>%
  mutate(dwc_countryCode = "BE")  %>%            
  mutate(dwc_continent = "Europe")
```

```{r}

occurrence %<>%
  
  mutate(dwc_decimalLongitude = round(dwc_decimalLongitude, digits = 5)) %>%
  mutate(dwc_decimalLatitude = round(dwc_decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(dwc_decimalLatitude = as.character(format(dwc_decimalLatitude, nsmall = 5))) %>%
   mutate(dwc_decimalLongitude = as.character(format(dwc_decimalLongitude, nsmall = 5)))
```
### coordinateUncertaintyInMeters

```{r}
occurrence %<>%
  mutate(dwc_coordinateUncertaintyInMeters = accuracy)
```


### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
#occurrence %<>% mutate(countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
occurrence %<>%
  mutate(dwc_municipality = location)
```

### verbatimcoordinates

### verbatimLatitude

### verbatimLongitude

### verbatimcoordinatesystem

```{r}
#occurrence %<>% mutate(verbatimcoordinatesystem = "Lambert coordinates") # 
```

### verbatimSRS

```{r}
#occurrence %<>% mutate(verbatimSRS = "Belgian Datum 1972")
```

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = Extracted)
```


### identificationVerificationStatus


```{r}
occurrence %<>% mutate(dwc_identificationVerificationStatus = case_when(occurrence$`validation status` == 'onbekend' ~'unknown'
                                                                      ,occurrence$`validation status` == 'goedgekeurd (automatische validatie)' ~ 'approved (automatic validation)'
                                                                      , TRUE ~ `validation status`  ))
```
### identificationRemarks


```{r}
occurrence %<>% mutate(dwc_identificationRemarks = `validator name`)
```



### scientificName 

```{r}
occurrence %<>% mutate(dwc_scientificName = `scientific name`)
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### scientificNameAuthorship

```{r}


```

### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = case_when(occurrence$`species type` == 'Soort' ~'species'
                                                 , TRUE ~ `species type` ))
```



### nomenclaturalCode

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICZN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(dwc_occurrenceStatus = case_when(dwc_individualCount > 0 ~ "Present",
                                        dwc_individualCount == 0 ~ "Absent",
                                        is.na(dwc_individualCount) ~ "Absent"
                                                  
                                       ))
head(occurrence, n = 5)
```



## Post-processing

```{r}

occurrence %<>% select(starts_with("dwc_"))
```



```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")

```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

Define the order of the output columns

```{r}
col_order <- c( "type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount","organismQuantity","organismQuantityType","occurrenceStatus","occurrenceRemarks"
               ,"samplingProtocol","samplingEffort"
               ,"eventDate","continent","countryCode","municipality"
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","kingdom","taxonRank"
               )
occurrence <- occurrence[ col_order]
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "np-profile-download", "data", "processed", "occurrence.csv"), na = "")
```


Visualize data on a map

```{r}
occurrenceMap <- occurrence
```

```{r}
occurrenceMap$decimalLatitude <- as.numeric(occurrenceMap$decimalLatitude)
```

```{r}
occurrenceMap$decimalLongitude <- as.numeric(occurrenceMap$decimalLongitude)
```

```{r}
head(occurrenceMap)
```



```{r}


head(occurrence)

mapview(occurrenceMap, xcol = "decimalLongitude", ycol = "decimalLatitude", crs = 4269, grid = FALSE)

```

