---
title: "Darwin Core Mapping UTM SQLite connection"
subtitle: "VBP-INBO-waarnemingen-occurrences_202_planten
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(tidyverse)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(sp)             # coordinate transformation
library(leaflet)        # coordinate transformation
library(widgetframe)    # coordinate transformation
library(sf)             # coordinate transformation
library(lubridate)      # for the date
library(ggplot2)        # for map
library(mapview)        # to view map
library(DBI)            # database stuff
library(RSQLite)        # sqlite connection
library(reader)
```

# Read source data

Create a data frame `input_data` from the source data:
The source data was corrected in Excel
Muskrat occurrences opened in openRefine
Obsolete columns removed
some columns renamed to DwC term
File exported to csv

#load and list files

# Verbind met SQLite

```{r eval=FALSE, include=FALSE}
con <- dbConnect(SQLite(), (here::here("observations-instituut-voor-natuur-en-bosonderzoek_all.sqlite" )))
```

# Bekijk alle tabellen én views

```{r eval=FALSE, include=FALSE}
dbListTables(con)  # views staan hier ook tussen
```

```{r eval=FALSE, include=FALSE}
input_interim <- dbReadTable(con, "INBO_WNMN_PLANTS_native")  # werkt ook voor views
```


# Resultaat bekijken

```{r}
head(input_interim)
```



```{r eval=FALSE, include=FALSE}
input_interim <- read.csv(here::here("datasets", "np-datavragen-data-vaatplanten", "data", "interim", "occurrence_planten_inbo_wnmn.csv" ))
```



# Resultaat bekijken

```{r}
head(input_interim)
```




Preview data:


# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_interim %<>% remove_empty("rows")
```

```{r}
input_interim %>% head(n = 5)
```

# embargo date

```{r}
input_interim <- input_interim %>%
  filter(is.na(embargo_date) | embargo_date < today())
```



# Georeference source data

# Zet om naar sf-object met WGS84 (verondersteld: decimalLongitude & decimalLatitude)
```{r}
sf_points <- st_as_sf(input_interim, coords = c("decimalLongitude", "decimalLatitude"), crs = 4326)
```

# Transformeer naar UTM zone 31N (voor België)
```{r}
sf_utm <- st_transform(sf_points, crs = 32631)

```
# check sfUTM

```{r}
head(sf_utm)
```



# Extraheer UTM x/y

```{r}

coords <- st_coordinates(sf_utm)
input_interim$utm_easting <- coords[,1]
input_interim$utm_northing <- coords[,2]
```

# Bereken centroïde van het 1 km-hok

```{r}

input_interim$utm_centroid_easting <- floor(input_interim$utm_easting / 1000) * 1000 + 500
input_interim$utm_centroid_northing <- floor(input_interim$utm_northing / 1000) * 1000 + 500
```

# Bekijk resultaat

```{r}

head(input_interim)
```

# Zet de UTM-coördinaten terug naar WGS84 (decimale graden)
```{r}
sf_wgs84 <- st_transform(sf_utm, crs = 4326)
```

```{r}
head(sf_wgs84)
```


# Haal de coördinaten op
```{r}

coords <- st_coordinates(sf_wgs84)
```

# Voeg de decimale graden toe aan je data frame
```{r}
input_interim$dwc_longitude <- coords[, 1]
input_interim$dwc_latitude <- coords[, 2]
```

# Resultaat bekijken

```{r}
head(input_interim)
```
# Maak een sf-object van de centroiden in UTM

```{r}

sf_centroids <- st_as_sf(input_interim,
                         coords = c("utm_centroid_easting", "utm_centroid_northing"),
                         crs = 32631, # UTM zone 31N
                         remove = FALSE)
```
# Zet om naar WGS84

```{r}

sf_centroids_wgs84 <- st_transform(sf_centroids, crs = 4326)
```

# Extraheer de WGS84-coördinaten
```{r}

centroid_coords <- st_coordinates(sf_centroids_wgs84)
```

# Voeg de WGS84 centroid-coördinaten toe aan je dataframe
```{r}
input_interim$dwc_centroid_longitude <- centroid_coords[, 1]
input_interim$dwc_centroid_latitude <- centroid_coords[, 2]
```



# Occurrence core

## Pre-processing

Create a dataframe occurrence data only 

```{r}
occurrence <- input_interim
```

## remove obsolete columns

```{r}
occurrence %<>% select(-c(is_escape, embargo_date))
```
## filter for species group

Define which group of species to map to DwC

```{r eval=FALSE, include=FALSE}
occurrence %<>%
        filter(`species_group` == 'Planten') 
        
head(occurrence, n = 5 )

```



# Term mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

Start with record-level terms which contain metadata about the dataset (which is generally the same for all records).

# Event

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/10.15468/h25ces")
```

### add row numbers to genrate occurrenceID unique


```{r}
occurrence %<>% mutate(ID_row = row_number())
```


### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en") # e.g. "en"
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/") 
# e.g. "http://creativecommons.org/publicdomain/zero/1.0/"
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "INBO") # e.g. "INBO"
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```
### accessRights

```{r}
occurrence %<>% mutate(dwc_references = references) 
```



### datasetID

```{r}
#occurrence %<>% mutate(datasetID = "insert doi") 
```

### institutionCode

```{r}
occurrence %<>% mutate(dwc_institutionCode = "INBO") # e.g. "INBO"
```
### institutionID

```{r}
occurrence %<>% mutate(dwc_institutionID = "https://ror.org/00j54wy13") # e.g. "INBO"
```
### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "INBO collaborator Occurrences of plants recorded in waarnemingen.be in Flanders, Belgium") # e.g. "Checklist of non-native freshwater fishes in Flanders, Belgium"
```

The following terms contain information about the taxon:

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### informationWithHeld

### dataGeneralizations

```{r}
occurrence %<>% mutate(dwc_dataGeneralizations = "coordinates generalized to 1km grid")
```



### occurrenceID

```{r}
occurrence %<>% mutate(dwc_occurrenceID = str_c("INBO:OCC:WNMN:", occurrenceID))
  
```

```{r}
occurrence %<>% mutate(dwc_catalogNumber = references)
```



### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy = recordedBy)
```

### identifiedBy

```{r}
occurrence %<>% mutate(dwc_identifiedBy = indentifiedBy)
```



### individualCount

```{r}
occurrence %<>%  mutate(dwc_individualCount = individualCount) 
                
```

```{r}
occurrence$individualCount %<>% replace_na(0)
```

### organismQuantity


```{r}
occurrence %<>%  mutate(dwc_organismQuantity = individualCount) 
                
```



### organismQuentityType

```{r}
occurrence %<>%  mutate(dwc_organismQuantityType = 'count estimate') 
                
```


### sex

```{r}
occurrence %<>% mutate(dwc_sex = case_when(occurrence$sex == 'U' ~'unknown'
                                       ,occurrence$sex == 'F' ~'female'
                                       ,occurrence$sex == 'M' ~'male',
                                       TRUE ~ sex))
```



### lifeStage

```{r}
occurrence %<>% mutate(dwc_lifestage = case_when(occurrence$`lifeStage` == 'onbekend' ~'unknown'
                                                 , TRUE ~ `lifeStage`))
```


### behavior

### occurrenceRemarks

```{r}
occurrence %<>% mutate(dwc_collectionCode = 'INBO-WNMN')
```

## samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingEffort = 'casual observation')
```

```{r}
occurrence %<>% rename(dwc_samplingProtocol = samplingProtocol)
 


```

### samplingEffort


### eventDate

eventDate already ok in source file (should be)

```{r}


#occurrence %<>% ymd(date)

occurrence %<>% mutate(dwc_eventDate = eventDate)
  


```

# Location

```{r}
occurrence %<>%
  mutate(dwc_decimalLongitude = dwc_centroid_longitude) %>%
  mutate(dwc_decimalLatitude = dwc_centroid_latitude) %>%
  mutate(dwc_geodeticDatum = "WGS84") %>%
  mutate(dwc_coordinateUncertaintyInMeters = "701") %>%
  mutate(dwc_stateProvince = stateProvince) %>%
  mutate(dwc_countryCode = "BE")  %>%            
  mutate(dwc_continent = "Europe")
```

```{r}

occurrence %<>%
  
  mutate(dwc_decimalLongitude = round(dwc_decimalLongitude, digits = 5)) %>%
  mutate(dwc_decimalLatitude = round(dwc_decimalLatitude, digits = 5))
```

```{r}
occurrence %<>%   
   mutate(dwc_decimalLatitude = as.character(format(dwc_decimalLatitude, nsmall = 5))) %>%
   mutate(dwc_decimalLongitude = as.character(format(dwc_decimalLongitude, nsmall = 5)))
```
### coordinateUncertaintyInMeters

```{r}
occurrence %<>%
  mutate(dwc_coordinateUncertaintyInMeters = '707')
```


### continent

```{r}
##occurrence %<>% mutate(continent = "Europe") # e.g. "Belgium = BE"
```

### countryCode

```{r}
#occurrence %<>% mutate(countryCode = "BE") # e.g. "Belgium = BE"
```

### municipality

municipality already in source file

```{r}
occurrence %<>% mutate(dwc_municipality = sub(" -.*","",locality))
```


```{r}
occurrence %<>%
  mutate(dwc_locality = locality)
```

### verbatimcoordinates


### verbatimLatitude

### verbatimLongitude

### verbatimcoordinatesystem

```{r}
occurrence %<>% mutate(verbatimcoordinatesystem = "UTM 1km") # 
```

### verbatimSRS

```{r}
#occurrence %<>% mutate(verbatimSRS = "Belgian Datum 1972")
```

### decimalLatitude

### decimalLongitude

### geodeticDatum

### coordinateUncertaintyInMeters

### georeferenceRemarks

### identifiedBy



### identificationVerificationStatus

```{r}
occurrence %<>% mutate(dwc_identificationVerificationStatus = identificationVerificationStatus)
```

```{r eval=FALSE, include=FALSE}
occurrence %<>% mutate(dwc_identificationVerificationStatus = case_when(occurrence$`validation status` == 'onbekend' ~'unknown'
                                                                      ,occurrence$`validation status` == 'goedgekeurd (automatische validatie)' ~ 'approved (automatic validation)'
                                                                      , TRUE ~ `validation status`  ))
```
### identificationRemarks

## dynamicProperties

```{r}
occurrence %<>% mutate(dwc_dynamicProperties = '{"rbac":"false","rbac_allowed":"HIGHRES"}')
```




### scientificName 

```{r}
occurrence %<>% mutate(dwc_scientificName = scientificName)
```


### vernacularName 

```{r}
occurrence %<>% mutate(dwc_vernacularName = vernacularName)
```


### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Plantae")
```


### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = taxonRank )
```



### nomenclaturalCode

```{r}
occurrence %<>% mutate(nomenclaturalCode = "ICBN") # e.g. "ICZN"
```

### occurrenceStatus

```{r}
occurrence %<>% 
    ##select(individualCount) %>%
    mutate(dwc_occurrenceStatus = case_when(dwc_individualCount > 0 ~ "Present",
                                        dwc_individualCount == 0 ~ "Absent",
                                        is.na(dwc_individualCount) ~ "Absent"
                                                  
                                       ))
head(occurrence, n = 5)
```



## Post-processing

```{r}

occurrence %<>% select(starts_with("dwc_"))
```



```{r}
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")

```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

Define the order of the output columns

```{r eval=FALSE, include=FALSE}
col_order <- c( "type","language","license","rightsHolder","accessRights","datasetID"
               ,"institutionCode","datasetName","basisOfRecord","occurrenceID","recordedBy"
               ,"individualCount","organismQuantity","organismQuantityType","occurrenceStatus","occurrenceRemarks"
               ,"samplingProtocol","samplingEffort"
               ,"eventDate","continent","countryCode"
               ,"decimalLatitude","decimalLongitude","geodeticDatum","coordinateUncertaintyInMeters"
               ,"identifiedBy","scientificName","kingdom","taxonRank"
               )
occurrence <- occurrence[ col_order]
```

Preview data:



```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("datasets", "np-datavragen-data-vaatplanten", "data", "processed", "occurrence_planten.csv"), na = "")
```


# Visualize data on a map OriginalCoördinates

```{r}
occurrenceMap <- occurrence %>%
 filter(identifiedBy == "Jeroen Vanden Borre")
```

```{r}
occurrenceMap$decimalLatitude <- as.numeric(occurrenceMap$decimalLatitude)
```

```{r}
occurrenceMap$decimalLongitude <- as.numeric(occurrenceMap$decimalLongitude)
```

```{r}
head(occurrenceMap)
```



```{r}


head(occurrence)

mapview(occurrenceMap, xcol = "decimalLongitude", ycol = "decimalLatitude", crs = 4269, grid = FALSE)

```

# Visualize data on a map UTMCoördinates

```{r}
occurrenceMapUTM <- occurrence %>% 
 filter(identifiedBy == "Jeroen Vanden Borre")
```

```{r}
occurrenceMapUTM$centroid_latitude <- as.numeric(occurrenceMapUTM$centroid_latitude)
```

```{r}
occurrenceMapUTM$centroid_longitude <- as.numeric(occurrenceMapUTM$centroid_longitude)
```

```{r}
head(occurrenceMapUTM)
```
```{r}
problems(input_interim)
```



```{r}


head(occurrence)

mapview(occurrenceMapUTM, xcol = "centroid_longitude", ycol = "centroid_latitude", crs = 4269, grid = FALSE)

```

